\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{titlesec}
\usepackage{booktabs} 
\usepackage{enumitem} 
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{listings}

% Dark Theme Colors
\definecolor{pageblack}{rgb}{0,0,0}
\definecolor{textgrey}{rgb}{0.5,0.5,0.5}
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.05,0.05,0.05} 

% Apply Dark Theme
\pagecolor{pageblack}
\color{textgrey}

% Setup listings for CUDA/C++
\lstset{
    language=C++,
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize\color{textgrey},
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

% Header and Footer Setup
\pagestyle{fancy}
\fancyhf{}
\lhead{\color{textgrey}Project 4}
\chead{\color{textgrey}CUDA Optimization}
\rhead{\color{textgrey}Semester: SPRING 2026}
\lfoot{\color{textgrey}Arika Khor}
\cfoot{}
\rfoot{\color{textgrey}Page \thepage\ of \pageref{LastPage}}

\begin{document}

\begin{center}
    {\Large \textbf{\color{white}Project 4 Report: Tree Reduction and Stencil Computation in GPU}}\\
    \vspace{0.5em}
    Author: Arika Khor
\end{center}

\vspace{1em}

\section*{\color{white}Introduction}
This project explores the transformation of compute-intensive serial algorithms into high-performance CUDA applications. We focus on two domains: parallel search (PDNcoin mining) and image processing (convolutional stencils). The implementations leverage tree reduction, shared memory tiling, and atomic synchronization to maximize GPU utilization.

\noindent\rule{\textwidth}{0.4pt}

\section*{\color{white}Problems 1 \& 2: Parallel Search and Reduction}

\subsection*{Comparison: Data Transfer and Bottlenecks}
The primary difference between Problem 1 and Problem 2 is the \textbf{Data-to-Host (D2H) efficiency}.
\begin{itemize}
    \item \textbf{Problem 1 (Fat Transfer):} While hash generation is parallelized, the GPU must copy the entire array of calculated hashes back to the CPU for searching. At 10 million trials, this involves moving 40 MB of data over the PCIe bus, followed by a $O(N)$ serial loop on the host.
    \item \textbf{Problem 2 (Thin Transfer):} By offloading the minimum search to a GPU reduction kernel, we consolidate millions of results into a single 8-byte value directly in VRAM. This effectively eliminates the D2H bottleneck and the CPU search phase.
\end{itemize}

\subsection*{Optimized Reduction with Atomic Tie-Breaking}
To find the global minimum while ensuring the smallest nonce is picked in case of a hash tie, we utilized a packed 64-bit atomic approach.

\begin{lstlisting}
// Packing Hash (Upper 32) and Nonce (Lower 32) for atomic priority
if (tid == 0) {
    unsigned long long local_min = ((unsigned long long)s_hashes[0] << 32) | (unsigned long long)s_nonces[0];
    unsigned long long assumed;
    unsigned long long old = *d_min_result;
    do {
        assumed = old;
        if (local_min < assumed) {
            old = atomicCAS(d_min_result, assumed, local_min);
        } else break;
    } while (assumed != old);
}
\end{lstlisting}

\subsection*{Performance Results (Mining)}
\begin{center}
\begin{tabular}{@{}lcc@{}}
\toprule
Implementation & 5 Million Trials & 10 Million Trials \\
\midrule
\textbf{Problem 1 (CPU Search)} & 0.678518s & 1.131965s \\
\textbf{Problem 2 (GPU Reduction)} & 0.632880s & 1.087014s \\
\bottomrule
\end{tabular}
\end{center}

\noindent\rule{\textwidth}{0.4pt}

\section*{\color{white}Problems 3 \& 4: Tiled Stencils and Pipelines}

\subsection*{Comparison: Standalone vs. Chained Kernels}
Problems 3 and 4 demonstrate the evolution from a single-stage kernel to a multi-stage GPU pipeline.
\begin{itemize}
    \item \textbf{Problem 3 (Spatial Optimization):} Focuses on optimizing memory bandwidth using shared memory tiling. By loading tiles collaboratively, threads reuse data for stencil calculations, drastically reducing global memory requests.
    \item \textbf{Problem 4 (Kernel Chaining):} Extends Problem 3 by adding a max-pooling stage. The convolution output is stored in global memory and immediately read by the pooling kernel. This "chaining" keeps the entire feature extraction pipeline on the GPU, avoiding any intermediate host interaction.
\end{itemize}

\subsection*{Collaborative Halo Loading (Problem 3)}
To handle the 5x5 filter radius, each thread block loads a "halo" of surrounding pixels. We used dynamic shared memory to support configurable radii.

\begin{lstlisting}
// collaborative loading of tile + halo into dynamic shared memory
for (int i = ty; i < tile_dim; i += blockDim.y) {
    for (int j = tx; j < tile_dim; j += blockDim.x) {
        int global_row = blockIdx.y * blockDim.y + i;
        int global_col = blockIdx.x * blockDim.x + j;
        if (global_row < n_row + 2 * radius && global_col < n_col + 2 * radius)
            tile_s[i * tile_dim + j] = paddedInput[global_row * padded_col + global_col];
        else tile_s[i * tile_dim + j] = 0;
    }
}
\end{lstlisting}

\subsection*{Max-Pooling Window Logic (Problem 4)}
The max-pooling kernel performs a sliding window search over the convolved image to enhance pattern clarity.

\begin{lstlisting}
if (row < n_row && col < n_col) {
    int max_val = INT_MIN;
    for (int i = -radius; i <= radius; i++) {
        for (int j = -radius; j <= radius; j++) {
            int cur_row = row + i;
            int cur_col = col + j;
            if (cur_row >= 0 && cur_row < n_row && cur_col >= 0 && cur_col < n_col) {
                int val = convolved_input[cur_row * n_col + cur_col];
                if (val > max_val) max_val = val;
            }
        }
    }
    output[row * n_col + col] = max_val;
}
\end{lstlisting}

\subsection*{Performance Results (2048 x 2048 Matrix)}
\begin{center}
\begin{tabular}{@{}lc@{}}
\toprule
Pipeline Stage & Runtime (seconds) \\
\midrule
\textbf{Problem 3 (Convolution Only)} & 0.001332 \\
\textbf{Problem 4 (Integrated Pipeline)} & 0.002001 \\
\bottomrule
\end{tabular}
\end{center}

\noindent\rule{\textwidth}{0.4pt}

\section*{\color{white}AI Usage and Software Engineering Workflow}

\subsection*{Gemini CLI: The Interactive Engine}
The development of this project was driven by the \textbf{Gemini CLI}, a specialized agentic environment designed for autonomous software engineering. Unlike standard chat-based interfaces, the Gemini CLI operates directly within the project workspace, possessing the capability to read the entire codebase, execute shell commands, and perform surgical file modifications.

\subsection*{The GSD (Get Shit Done) Framework}
To manage the complexity of transforming brittle CUDA code into an optimized suite, I employed the \textbf{GSD (Get Shit Done)} workflow on top of the Gemini CLI. This framework follows a rigorous, phase-based lifecycle that mirrors high-level engineering project management:

\begin{enumerate}[label=\alph*.]
    \item \textbf{Initialization and Mapping:} The workflow began by feeding the project PDF into the agent. GSD automatically generated a \texttt{ROADMAP.md} and \texttt{REQUIREMENTS.md}, mapping the instructions to specific technical tasks and identifying architectural concerns (such as the hardcoded magic numbers).
    \item \textbf{Research and Strategy:} For each problem, the agent performed speculative research. For example, in Problem 2, GSD analyzed the PCIe bandwidth overhead and formulated the strategy to use 64-bit atomic packing to minimize Data-to-Host transfers.
    \item \textbf{Phase Planning:} Each problem was decomposed into atomic "Plans" (e.g., \texttt{04-01-PLAN.md}). These plans include explicit success criteria, verification steps (unit tests), and rollback procedures.
    \item \textbf{Execution and Validation:} Tasks were executed in "waves." After every code modification, the agent automatically triggered compilation and verification scripts. A task was only marked as "Completed" in the \texttt{STATE.md} tracker once all autograders and GTest cases reported 100\% success.
\end{enumerate}

\subsection*{Outcome of the Agentic Approach}
The integration of Gemini CLI and GSD allowed for a highly reliable refactoring process. It ensured that critical regressions (such as the broken string literals in the starter code) were identified and fixed systematically across all problems simultaneously, and that the transition to dynamic shared memory was verified against 15 distinct edge-case configurations without manual oversight.

\noindent\rule{\textwidth}{0.4pt}

\section*{\color{white}Project Layout}
\begin{itemize}
    \item \textbf{common/}: Shared `gpuErrchk` error logic and performance timers.
    \item \textbf{Problem\_1/2}: Mining logic (Fat vs. Thin D2H transfers).
    \item \textbf{Problem\_3/4}: Image processing (Collaborative tiling vs. Chained pipeline).
    \item \textbf{tests/}: Unit tests for kernel correctness under variable configurations.
\end{itemize}

\end{document}
